---
title: "Conformal"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr) 
library(ggplot2)
library(knitr)
library(broom)
library(tidyr)
library(progress)
library(pbapply)
pboptions(type='none')
library(dbscan)
library(gridExtra)
library(roahd)
library(conformalInference)
library(mgcv)
library(rgl)
```

```{r}
table(data$Gender, data$Smoking.status)
```


###Conformal prediction on Sleep duration when gender= female and smoking=1 (61 osservazioni)
```{r}
dt <- data[data$Gender == "Female" & data$Smoking.status == 1, c(2,6,12,13,15)]
```


```{r}
covariate_names <- c("Age", "Caffeine.consumption", "Alcohol.consumption", "Exercise.frequency")
y <- dt$Sleep.duration
x <- dt[, covariate_names]

attach(dt)
```

```{r}
median_Caffeine <- median(dt$Caffeine.consumption, na.rm = TRUE)
median_Alcohol <- median(dt$Alcohol.consumption, na.rm = TRUE)
median_Exercise <- median(dt$Exercise.frequency, na.rm = TRUE)

# Age has different values
age_grid <- seq(from = min(dt$Age, na.rm = TRUE), to = max(dt$Age, na.rm = TRUE), length.out = 100)

# Test grid
test_grid <- data.frame(
  Age = age_grid,
  Caffeine.consumption = rep(median_Caffeine, length(age_grid)),
  Alcohol.consumption = rep(median_Alcohol, length(age_grid)),
  Exercise.frequency = rep(median_Exercise, length(age_grid))
)

```

```{r}
model_gam <- gam(y ~ s(Age, bs = 'cr') + Caffeine.consumption + Alcohol.consumption + Exercise.frequency, data = dt)

pred=predict(model_gam,newdata=test_grid)
plot(age_grid,pred)
```

###Training function
```{r}
train_gam <- function(x, y, out = NULL) {
  colnames(x) <- c('var1','var2','var3','var4')
  train_data <- data.frame(y, x)
  
  # GAM model
  model_gam <- gam(y ~ s(var1, bs = 'cr') + var2 + var3 + var4, data = train_data)
}
```

###Prediction function
```{r}
predict_gam <- function(obj, new_x) {
  new_x <- data.frame(new_x)
  colnames(new_x) <- c('var1','var2','var3','var4')

  predict.gam(obj, new_x)
}
```


###Full conformal framework
```{r}
alpha <- 0.1

result_full <- data.frame(Age = numeric(), Prediction = numeric(), Lower = numeric(), Upper = numeric())

for(i in 1:nrow(test_grid)){
  predictions <- conformal.pred(
    x = cbind(Age,Caffeine.consumption,Alcohol.consumption,Exercise.frequency),
    y = Sleep.duration,
    x0 = c(age_grid[i],median_Caffeine,median_Alcohol,median_Exercise),
    alpha = alpha,
    verbose = FALSE,
    train.fun = train_gam,
    predict.fun = predict_gam,
    num.grid.pts = 200
  )
  result_full <- rbind(result_full, data.frame(
    Age = age_grid[i],
    Prediction = predictions$pred,
    Lower = predictions$lo,
    Upper = predictions$up
  ))
}

head(result_full)
#saveRDS(result_full, "result_full_F1.rds")

```


```{r}
ggplot(result_full, aes(x = Age, y = Prediction)) +
  geom_line(color = "blue", size = 1) +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), fill = "lightblue", alpha = 0.4) +
  labs(title = "Sleep duration with Full Conformal Prediction",
       x = "Age",
       y = "Sleep duration") +
  theme_minimal()
```


###Split conformal framework
```{r}
alpha <- 0.1

result_split <- data.frame(Age = numeric(), Prediction = numeric(), Lower = numeric(), Upper = numeric())

for(i in 1:nrow(test_grid)){
  predictions <- conformal.pred.split(
    x = cbind(Age,Caffeine.consumption,Alcohol.consumption,Exercise.frequency),
    y = Sleep.duration,
    x0 = c(age_grid[i],median_Caffeine,median_Alcohol,median_Exercise),
    alpha = alpha,
    verbose = FALSE,
    train.fun = train_gam,
    predict.fun = predict_gam,
    rho=0.5
  )
  result_split <- rbind(result_split, data.frame(
    Age = age_grid[i],
    Prediction = predictions$pred,
    Lower = predictions$lo,
    Upper = predictions$up
  ))
}

head(result_split)
#saveRDS(result_split, "result_split_F1.rds")
```


```{r}
ggplot(result_split, aes(x = Age, y = Prediction)) +
  geom_line(color = "blue", size = 1) +  
  geom_ribbon(aes(ymin = Lower, ymax = Upper), fill = "lightblue", alpha = 0.4) +
  labs(title = "Sleep duration with Split Conformal Prediction",
       x = "Age",
       y = "Sleep duration") +
  theme_minimal()
```

```{r}
detach(dt)
```


###Conformal prediction on Sleep duration when gender= male and smoking=1  (93 osservazioni)
```{r}
dt <- data[data$Gender == "Male" & data$Smoking.status == 1, c(2,6,12,13,15)]

saveRDS(result_full, "result_full_M1.rds")
saveRDS(result_split, "result_split_M1.rds")
```


###Conformal prediction on Sleep duration when gender= female and smoking=0 (163 osservazioni)
```{r}
dt <- data[data$Gender == "Female" & data$Smoking.status == 0, c(2,6,12,13,15)]

saveRDS(result_full, "result_full_F0.rds")
saveRDS(result_split, "result_split_F0.rds")
```

###Conformal prediction on Sleep duration when gender= male and smoking=0 (135 osservazioni)
```{r}
dt <- data[data$Gender == "Male" & data$Smoking.status == 0, c(2,6,12,13,15)]

saveRDS(result_full, "result_full_F1.rds")
saveRDS(result_split, "result_split_F1.rds")
```




### - Try with different values for caffeine when gender=female and smoking=0
```{r}
dt <- data[data$Gender == "Female" & data$Smoking.status == 0, c(2,6,12,13,15)]
```


```{r}
covariate_names <- c("Age", "Caffeine.consumption", "Alcohol.consumption", "Exercise.frequency")
y <- dt$Sleep.duration
x <- dt[, covariate_names]

attach(dt)
```


```{r}
median_Alcohol <- median(dt$Alcohol.consumption, na.rm = TRUE)
median_Exercise <- median(dt$Exercise.frequency, na.rm = TRUE)
median_Age <- median(dt$Age, na.rm = TRUE)

# Age has different values
caffeine_grid <- seq(from = min(dt$Caffeine.consumption, na.rm = TRUE), to = max(dt$Caffeine.consumption, na.rm = TRUE), length.out = 100)

# Test grid
test_grid <- data.frame(
  Age = rep(median_Age, length(caffeine_grid)),
  Caffeine.consumption = caffeine_grid,
  Alcohol.consumption = rep(median_Alcohol, length(caffeine_grid)),
  Exercise.frequency = rep(median_Exercise, length(caffeine_grid))
)
```


```{r}
model_gam <- gam(y ~ Age + Caffeine.consumption + Alcohol.consumption + Exercise.frequency, data = dt)

pred=predict(model_gam,newdata=test_grid)

plot(Caffeine.consumption, y, pch = 16, col = "blue", xlab = "Caffeine consumption", ylab = "Response variable")

lines(caffeine_grid, pred, col = "red", lwd = 2)
```

###Training function
```{r}
train_gam <- function(x, y, out = NULL) {
  colnames(x) <- c('var1','var2','var3','var4')
  train_data <- data.frame(y, x)
  
  # GAM model
  model_gam <- gam(y ~ var1 + var2 + var3 + var4, data = train_data)
}
```

###Prediction function
```{r}
predict_gam <- function(obj, new_x) {
  new_x <- data.frame(new_x)
  colnames(new_x) <- c('var1','var2','var3','var4')

  predict.gam(obj, new_x)
}
```


###Full conformal framework
```{r}
alpha <- 0.1

result_full <- data.frame(Age = numeric(), Prediction = numeric(), Lower = numeric(), Upper = numeric())

for(i in 1:nrow(test_grid)){
  predictions <- conformal.pred(
    x = cbind(Age,Caffeine.consumption,Alcohol.consumption,Exercise.frequency),
    y = Sleep.duration,
    x0 = c(median_Age,caffeine_grid[i],median_Alcohol,median_Exercise),
    alpha = alpha,
    verbose = FALSE,
    train.fun = train_gam,
    predict.fun = predict_gam,
    num.grid.pts = 200
  )
  result_full <- rbind(result_full, data.frame(
    Caffeine = caffeine_grid[i],
    Prediction = predictions$pred,
    Lower = predictions$lo,
    Upper = predictions$up
  ))
}

head(result_full)
#saveRDS(result_full, "result_full_F1.rds")

```


```{r}
ggplot(result_full, aes(x = Caffeine, y = Prediction)) +
  geom_line(color = "blue", size = 1) +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), fill = "lightblue", alpha = 0.4) +
  labs(title = "Sleep duration with Full Conformal Prediction",
       x = "Caffeine",
       y = "Sleep duration") +
  theme_minimal()
```


### Performance
```{r}
evaluate_predictions <- function(x.test, y.test, x.grid, lower, upper, view_plot=T){
  covered <- (y.test >= lower)*(y.test<=upper)
  coverage <- mean(covered)
  width <- mean(upper-lower)
  
  if(view_plot){
    idx.sort <- sort(x.test$x, index.return=TRUE)$ix
    plot(x.test$x[idx.sort], y.test[idx.sort], col='lightblue', main=paste0("Prediction interval, alpha=",alpha, ", coverage=", round(coverage,2), ", width=", round(width,2)),
     xlab="x test", ylab="y test")
    lines(x.test$x[idx.sort], lower[idx.sort], lty=3, col='black', lwd=2)
    lines(x.test$x[idx.sort], upper[idx.sort], lty=3, col='black', lwd=2)
  }
  
  out <- list("coverage"=coverage,
              "width"=width)
  return(out)
}
```


```{r}
performance <- evaluate_predictions(
  x.test=data.frame("x" = dt$Age),
  y.test=dt$Sleep.duration,
  lower=result_full_F1$Lower,
  upper=result_full_F1$Upper)

print(performance)
# coverage 0.93
# width 3.932161
```


###Conformal prediction on Sleep efficiency when gender= female and smoking=1 (61 osservazioni)
```{r}
dt <- data[data$Gender == "Female" & data$Smoking.status == 1, c(2,7,12,13,15)]
```


```{r}
covariate_names <- c("Age", "Caffeine.consumption", "Alcohol.consumption", "Exercise.frequency")
y <- dt$Sleep.efficiency
x <- dt[, covariate_names]

attach(dt)
```

```{r}
median_Caffeine <- median(dt$Caffeine.consumption, na.rm = TRUE)
median_Alcohol <- median(dt$Alcohol.consumption, na.rm = TRUE)
median_Exercise <- median(dt$Exercise.frequency, na.rm = TRUE)

# Age has different values
age_grid <- seq(from = min(dt$Age, na.rm = TRUE), to = max(dt$Age, na.rm = TRUE), length.out = 100)

# Test grid
test_grid <- data.frame(
  Age = age_grid,
  Caffeine.consumption = rep(median_Caffeine, length(age_grid)),
  Alcohol.consumption = rep(median_Alcohol, length(age_grid)),
  Exercise.frequency = rep(median_Exercise, length(age_grid))
)

```

```{r}
model_gam <- gam(y ~ s(Age, bs = 'cr') + Caffeine.consumption + Alcohol.consumption + Exercise.frequency, data = dt)

pred=predict(model_gam,newdata=test_grid)

plot(Age, y, pch = 16, col = "blue", xlab = "Age", ylab = "Response variable")

lines(age_grid, pred, col = "red", lwd = 2)
#plot(age_grid, pred, col = "red", lwd = 2)
```

###Training function
```{r}
train_gam <- function(x, y, out = NULL) {
  colnames(x) <- c('var1','var2','var3','var4')
  train_data <- data.frame(y, x)
  
  # GAM model
  model_gam <- gam(y ~ s(var1, bs = 'cr') + var2 + var3 + var4, data = train_data)
}
```

###Prediction function
```{r}
predict_gam <- function(obj, new_x) {
  new_x <- data.frame(new_x)
  colnames(new_x) <- c('var1','var2','var3','var4')

  predict.gam(obj, new_x)
}
```


###Full conformal framework
```{r}
alpha <- 0.1

result_full <- data.frame(Age = numeric(), Prediction = numeric(), Lower = numeric(), Upper = numeric())

for(i in 1:nrow(test_grid)){
  predictions <- conformal.pred(
    x = cbind(Age,Caffeine.consumption,Alcohol.consumption,Exercise.frequency),
    y = Sleep.efficiency,
    x0 = c(age_grid[i],median_Caffeine,median_Alcohol,median_Exercise),
    alpha = alpha,
    verbose = FALSE,
    train.fun = train_gam,
    predict.fun = predict_gam,
    num.grid.pts = 200
  )
  result_full <- rbind(result_full, data.frame(
    Age = age_grid[i],
    Prediction = predictions$pred,
    Lower = predictions$lo,
    Upper = predictions$up
  ))
}

head(result_full)
#saveRDS(result_full, "result_full_F1.rds")

```


```{r}
ggplot(result_full, aes(x = Age, y = Prediction)) +
  geom_line(color = "blue", size = 1) +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), fill = "lightblue", alpha = 0.4) +
  labs(title = "Sleep efficiency with Full Conformal Prediction",
       x = "Age",
       y = "Sleep duration") +
  theme_minimal()
```


