
## Gender 
```{r}
table(data$Gender)
data$AgeGroup <- cut(data$Age,
                 breaks = seq(min(data$Age), max(data$Age), length.out = nclass.FD(data$Age)),                  include.lowest = T, right = T) 
table(data$AgeGroup, data$Gender)
boxplot(data$Age~data$Gender, col=c("pink", "blue"))
```
balanced subdivision, but different sizes in Age_group (e.g. males are older, only female kids)

## Smoking status
```{r}
table(data$Smoking.status)
table(data$AgeGroup, data$Smoking.status)
boxplot(data$Age~data$Smoking.status)
boxplot(data$Exercise.frequency~data$Smoking.status, col=c("white", "grey"))
```
smokers are ~1/3 of the population

## Exercise frequency
```{r}
table(data$Exercise.frequency)
table(data$AgeGroup, data$Exercise.frequency)
boxplot(data$Age~data$Exercise.frequency)
table(data$Gender, data$Exercise.frequency)
boxplot(data$Exercise.frequency~data$Gender, col=c("pink", "blue"))
```
30-50 years old people is the most sporty; it seems that women are less sporty than men, but with higher frequencies

## Caffeine consumption
```{r}
table(data$Caffeine.consumption)
table(data$AgeGroup, data$Caffeine.consumption)
table(data$Gender, data$Caffeine.consumption)
boxplot(data$Age~data$Caffeine.consumption)
boxplot(data$Caffeine.consumption~data$Gender)
colori <- ifelse(data$Gender == "Male", "blue", "pink")
plot(data$Caffeine.consumption, data$Age, col=colori)
```

## Alcohol consumption
```{r}
table(data$Alcohol.consumption)
table(data$AgeGroup, data$Alcohol.consumption)
table(data$Gender, data$Alcohol.consumption)
boxplot(data$Age~data$Alcohol.consumption)
boxplot(data$Alcohol.consumption~data$Gender)
colori <- ifelse(data$Gender == "Male", "blue", "pink")
plot(data$Alcohol.consumption, data$Age, col=colori)
```
less variability w.r.t. caffeine

##### Preliminary graphs: boxplot, hystograms
```{r}
fboxplot <- function (variabile, categorica=NULL, data){
  par(mfrow=c(1,1))
  if (is.null(categorica)){
    print(boxplot(data[[variabile]], main=variabile))
  } else {
    print(boxplot(data[[variabile]]~data[[categorica]], main=variabile, xlab=categorica, ylab=" "))
  }
}
```

```{r}
fistogramma <- function (variabile, categorica=NULL, data){
  if (is.null(categorica)){
    par(mfrow=c(1,1))
    print(hist(data[[variabile]], prob=TRUE, main=paste("istogramma di", variabile), xlab=" "))
  } else {
    m<-unique(data[[categorica]])
    title= paste("istogramma di", variabile, "rispetto a ", categorica)
    par(mfrow=c(1,length(m)))
    for (i in 1:length(m))
      print(hist(data[[variabile]][which(data[[categorica]]==m[i])], xlab= categorica))
    
  }
}
```


Some boxplot/histograms
```{r}
boxplot(datastd)
fboxplot("Age", data = data)
fboxplot("Age", "Gender", data = data)
fboxplot("Sleep.duration", "Gender", data)
fboxplot("Sleep.duration", "AgeGroup", data)
fistogramma("Sleep.duration", data = data)
fistogramma("Age", "Gender",data=data)
```


##### Pairs: coupled graphs with categorical variable
```{r}
pairs(data[,numeric_cols])
colori <- ifelse(data$Gender == "Male", "blue", "pink")
pairs(data[,numeric_cols], col=colori)

colori <- ifelse(data$Deep.sleep.percentage>41, "red", "green")
pairs(data[,c(2,6:13,15:p)], col=c("red","green"))

colori_gruppi <- rainbow(length(levels(data$AgeGroup)))
colori <- colori_gruppi[as.numeric(data$AgeGroup)] 
pairs(data[,numeric_cols], col=colori_gruppi)
```

##### Deep sleep vs Bedtime
```{r}
ggplot(data, aes(x = Bedtime, y = Deep.sleep.percentage)) +
  geom_point(alpha = 0.6, color = "blue") + 
  geom_smooth(method = "lm", color = "red", se = FALSE) + 
  labs(
    title = "Deep sleep vs Bedtime",
    x = "Bedtime hour",
    y = "Percentage of deep sleep"
  ) +
  theme_minimal()
```
It seems that as bedtime increases, deep sleep decreases

##### Deep sleep vs Wakeup time
```{r}
ggplot(data, aes(x = Wakeup.time, y = Deep.sleep.percentage)) +
  geom_point(alpha = 0.6, color = "green") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(
    title = " Deep sleep vs Wakeup time",
    x = "Wakeup hour",
    y = "Percentage of deep sleep"
  ) +
  theme_minimal()
```
It seems that as wakeup increases, deep sleep decreases (wakeup and bedtime are correlated)

