#### NON PARAMETRIC INFERENCE

##### Setting B and seed
```{r}
seed <- 482990
B <- 5000
```

##### Mann Whitney Univariate
```{r}
mann_whitney_analysis <- function(data, variabile, split_variable, split_criteria) {
  data1 <- data %>% filter(!!sym(split_variable) == split_criteria[1])
  data2 <- data %>% filter(!!sym(split_variable) == split_criteria[2])
  x1 <- data1[[variabile]]
  x2 <- data2[[variabile]]
  
  # Compute maximum to compare hystograms
 max_density <- max(    hist(x1, plot = FALSE)$density,     hist(x2, plot = FALSE)$density)
  
  # Istogrammi
  par(mfrow = c(1, 2))
  hist(x1,        main = paste(variabile, split_criteria[1]),        prob = T, 
       col = "pink",        ylim = c(0, max_density)) 
 hist(x2,        main = paste(variabile, split_criteria[2]),        prob = T, 
       col = "blue",        ylim = c(0, max_density))
  
  t1 <- as.data.frame(x1)
  colnames(t1) <- variabile
  t2 <- as.data.frame(x2)
  colnames(t2) <- variabile
  
  # DD-Plot
  ddp<-ddPlot(x = t1, y = t2, depth_params = list(method = 'Tukey'))
  print(ddp)
  
  # Mann-Whitney Test
  alpha <- 0.05
  mann_whitney_test <- wilcox.test(x1, x2, paired = FALSE, conf.level = 1 - alpha)
  p_value <- mann_whitney_test$p.value
  
  # Test results
  cat("Mann-Whitney Test Results for", variabile, "between", 
      split_criteria[1], "and", split_criteria[2], "\n")
  print(mann_whitney_test)
  
  return(p_value)
}
```


###### - Gender
```{r}
shapiro.test(filter(data,Gender=="Female")$Sleep.efficiency)
qqnorm(filter(data,Gender=="Female")$Sleep.efficiency)
```
not normal

###### - Sleep efficiency
```{r}
mann_whitney_analysis(data = data, variabile = "Sleep.efficiency", 
  split_variable = "Gender", split_criteria = c("Female", "Male"))
```
high pvalue (0.691607), no difference in distribution

###### - Sleep duration
```{r}
mann_whitney_analysis(data = data, variabile = "Sleep.duration", 
  split_variable = "Gender", split_criteria = c("Female", "Male")) 
```
high pvalue (0.423791), no difference in distribution


###### - Awakenings
```{r}
mann_whitney_analysis(data = data, variabile = "Awakenings", 
  split_variable = "Gender", split_criteria = c("Female", "Male")) 
```
pvalue=0.07668181, no difference in distribution


###### Young vs older people
```{r}
data <- data %>%
  mutate(young = ifelse(Age < 45, "Giovani", "Vecchi"))

shapiro.test(filter(data,Age<45)$Sleep.efficiency)
qqnorm(filter(data,Age<45)$Sleep.efficiency)
```
not normal

###### - Sleep efficiency
```{r}
mann_whitney_analysis(data = data, variabile = "Sleep.efficiency", 
  split_variable = "young", split_criteria = c("Giovani", "Vecchi")) 
```
pvalue=0.1079899

###### - Sleep duration
```{r}
mann_whitney_analysis(data = data, variabile = "Sleep.duration", 
  split_variable = "young", split_criteria = c("Giovani", "Vecchi"))
```
pvalue 0.06516022




###### Smokers VS Nonsmokers
```{r}
shapiro.test(filter(data,Smoking.status=="0")$Sleep.efficiency)
qqnorm(filter(data,Smoking.status=="0")$Sleep.efficiency)
```
not normal

###### - Sleep efficiency
```{r}
mann_whitney_analysis(data = data, variabile = "Sleep.efficiency", 
  split_variable = "Smoking.status", split_criteria = c("0", "1")) 
```
pvalue=5.954699e-07


###### - Sleep duration
```{r}
mann_whitney_analysis(
  data = data, variabile = "Sleep.duration", 
  split_variable = "Smoking.status", split_criteria = c("0", "1")) 
```
pvalue=0.5539101



###### High vs low caffeine
```{r}
data <- data %>%
  mutate(caffe = ifelse(Caffeine.consumption < 26, "nocaffe", "yescaffe"))
```


###### - Sleep duration
```{r}
mann_whitney_analysis(
  data = data, variabile = "Sleep.duration", 
  split_variable = "caffe", split_criteria = c("nocaffe", "yescaffe"))
```
pvalue=0.2778851


###### - Deep sleep percentage
```{r}
mann_whitney_analysis(
  data = data, variabile = "Deep.sleep.percentage", 
  split_variable = "caffe", split_criteria = c("nocaffe", "yescaffe")) 
```
pvalue 0.06



##### Permutational Multivariate Test (PMT)
Use this function with standardized data!
```{r}
permutation_test <- function(data, variables, split_variable, split_criteria, B , seed) {
  if (length(split_criteria) == 1 && grepl("<|>|=", split_criteria)) {
    condition <- paste0(split_variable, split_criteria)
    data1 <- dplyr::filter(data, eval(parse(text = condition))) %>%
      dplyr::select(all_of(variables))
    
    opposite_condition <- paste0("!", condition)
    data2 <- dplyr::filter(data, eval(parse(text = opposite_condition))) %>%
      dplyr::select(all_of(variables))
  } else {
    data1 <- dplyr::filter(data, !!sym(split_variable) == split_criteria[1]) %>%
      dplyr::select(all_of(variables))
    data2 <- dplyr::filter(data, !!sym(split_variable) == split_criteria[2]) %>%
      dplyr::select(all_of(variables))
  }
  
  n1 <- nrow(data1)
  n2 <- nrow(data2)
  n <- n1 + n2
  
  mean1 <- colMeans(data1)
  mean2 <- colMeans(data2)
  T20 <- as.numeric(t(mean1 - mean2) %*% (mean1 - mean2))
  
  T2 <- numeric(B)
  set.seed(seed)
  
  # Permutations
  for (perm in 1:B) {
    combined <- rbind(data1, data2)
    permutation <- sample(n)
    permuted <- combined[permutation, ]
    perm_data1 <- permuted[1:n1, ]
    perm_data2 <- permuted[(n1 + 1):n, ]
    
    mean1_perm <- colMeans(perm_data1)
    mean2_perm <- colMeans(perm_data2)
    T2[perm] <- t(mean1_perm - mean2_perm) %*% (mean1_perm - mean2_perm)
  }
  
  # Graphs
  par(mfrow = c(1, 2))
  
  hist(T2, xlim = range(c(T2, T20)), main = "Distribuzione permutazionale T2", xlab = "T2")
  abline(v = T20, col = "red", lwd = 2, lty = 2)
  
  # ECDF
  plot(ecdf(T2), main = "ECDF della distribuzione permutazionale", xlab = "T2", ylab = "F(T2)")
  abline(v = T20, col = "red", lwd = 2, lty = 2)
  
  # p-value
  p_val <- sum(T2 >= T20) / B
  
  cat("Test permutazionale multivariato\n")
  cat("Statistica T20:", T20, "\n")
  cat("p-value:", p_val, "\n")
  

  return(p_val)
}
```




###### Smoking status

###### -> Sleep efficiency + Sleep duration + Awakenings
```{r}
permutation_test(data = datastd,
  variables = c("Sleep.duration", "Sleep.efficiency", "Awakenings"),
  split_variable = "Smoking.status", split_criteria = levels(data$Smoking.status),
  B = B, seed = seed)
```
differences in sleep between smokers and non-smokers


###### High vs low caffeine intake

###### ->Sleep efficiency + Sleep duration + Awakenings
```{r}
datastd$aux<-ifelse(data$Caffeine.consumption <=24, "inf", "sup")
permutation_test(data = datastd,
  variables = c("Sleep.duration", "Sleep.efficiency", "Awakenings"),
  split_variable = "aux", split_criteria = c("inf","sup"),
  B = B, seed = seed)
```
0.41


###### Older vs younger people

###### -> Sleep efficiency + Sleep duration + Awakenings

```{r}
datastd$aux<-ifelse(data$Age <=35, "inf", "sup")
permutation_test(data = datastd,
  variables = c("Sleep.duration", "Sleep.efficiency", "Awakenings"),
  split_variable = "aux", split_criteria = c("inf","sup"),
  B = B, seed = seed)
```
0.0998


##### Exercise frequency
```{r}
datastd$aux<-ifelse(data$Exercise.frequency==0, "inf", "sup")
permutation_test(data = datastd,
  variables = c("Sleep.duration", "Sleep.efficiency", "Awakenings"),
  split_variable = "aux", split_criteria = c("inf","sup"),
  B = B, seed = seed)
```



##### Alcohol assumption
```{r}
datastd$aux<-ifelse(data$Alcohol.consumption==0, "inf", "sup")
permutation_test(data = datastd,
  variables = c("Sleep.duration", "Sleep.efficiency", "Awakenings"),
  split_variable = "aux", split_criteria = c("inf","sup"),
  B = B, seed = seed)
```



###### Gender

###### -> Sleep efficiency + Sleep duration + Awakenings
```{r}
permutation_test(data = datastd,
  variables = c("Sleep.duration", "Sleep.efficiency", "Awakenings"),
  split_variable = "Gender", split_criteria = c("Female", "Male"),
  B = B, seed = seed)
```
pvalue 0.26 

###### -> Deep sleep percentage + Sleep duration + Awakenings
```{r}
permutation_test(data = datastd,
  variables = c("Deep.sleep.percentage", "Sleep.duration", "Awakenings"),
  split_variable = "Gender", split_criteria = c("Female", "Male"),
  B = B, seed = seed)
```
p-value 0.09


###### - Deep sleep percentage graphs
```{r}
mann_whitney_analysis(data = datajitt, variabile = "Deep.sleep.percentage", 
  split_variable = "Gender", split_criteria = c("Female", "Male")) 
```

```{r}
mann_whitney_analysis(data = datajitt, variabile = "Deep.sleep.percentage", 
  split_variable = "Smoking.status", split_criteria = c("0", "1")) 
```

```{r}
datajitt <- datajitt %>%
  mutate(young = ifelse(Age < 45, "Giovani", "Vecchi"))

mann_whitney_analysis(data = datajitt, variabile = "Deep.sleep.percentage", 
  split_variable = "young", split_criteria = c("Giovani", "Vecchi")) 
```

```{r}
datajitt <- datajitt %>%
  mutate(exercise = ifelse(Exercise.frequency <=0, "noesercizio", "siesercizio"))

mann_whitney_analysis(data = datajitt, variabile = "Deep.sleep.percentage", 
  split_variable = "exercise", split_criteria = c("noesercizio", "siesercizio")) 
```

```{r}
datajitt <- datajitt %>%
  mutate(caffe = ifelse(Exercise.frequency <=25, "pococaffe", "moltocaffe"))

mann_whitney_analysis(data = datajitt, variabile = "Deep.sleep.percentage", 
  split_variable = "caffe", split_criteria = c("pococaffe", "moltocaffe")) 
```


####ANOVA PERMUTATION

```{r}
#fully parametric (and admittedly not robust) way
fit <- aov(data$Sleep.efficiency ~ data$Smoking.status)
summary(fit)

#permutational ANOVA
T0 <- summary(fit)[[1]][1,4] # extract the test statistic
T0
B<-1000
T_stat <- numeric(B)
n <- dim(data)[1]
for(perm in 1:B){
# Permutation:
permutation <- sample(1:n)
data_perm <- data$Sleep.efficiency[permutation]
fit_perm <- aov(data_perm ~ data$Smoking.status)
# Test statistic:
T_stat[perm] <- summary(fit_perm)[[1]][1,4]
}
hist(T_stat,xlim=range(c(T_stat,T0)),breaks=30)
abline(v=T0,col=3,lwd=2)
plot(ecdf(T_stat),xlim=c(-1,20))
abline(v=T0,col=3,lwd=4)
p_val <- sum(T_stat>=T0)/B
p_val
```


```{r}
exercise<-as.factor(data$Exercise.frequency)
fit <- aov(data$Sleep.efficiency ~ exercise)
summary(fit)

#permutational ANOVA
T0 <- summary(fit)[[1]][1,4] # extract the test statistic
T0
B<-1000
T_stat <- numeric(B)
n <- dim(data)[1]
for(perm in 1:B){
# Permutation:
permutation <- sample(1:n)
data_perm <- data$Sleep.efficiency[permutation]
fit_perm <- aov(data_perm ~ exercise)
# Test statistic:
T_stat[perm] <- summary(fit_perm)[[1]][1,4]
}
hist(T_stat,xlim=range(c(T_stat,T0)),breaks=30)
abline(v=T0,col=3,lwd=2)
plot(ecdf(T_stat),xlim=c(-1,20))
abline(v=T0,col=3,lwd=4)
p_val <- sum(T_stat>=T0)/B
p_val


```

#permutation two-way anova
```{r}
summary(aov(data$Sleep.efficiency ~ data$Smoking.status + exercise + data$Smoking.status:exercise))

#test first if the interaction is significant
T0_int <- summary.aov(aov(data$Sleep.efficiency ~ data$Smoking.status + exercise + data$Smoking.status:exercise))[[1]][3,4] # extract the test statistic
T0_int
aov.H0int <- aov(data$Sleep.efficiency ~ data$Smoking.status + exercise)
residuals.H0int  <- aov.H0int$residuals

T_int <- numeric(B)
for(perm in 1:B){
permutation <- sample(n)
# permute the residuals
residuals.H0int <- residuals.H0int[permutation]
eff.perm.H0int <- aov.H0int$fitted + residuals.H0int
# re-fit full model to calculate statistic
T_int[perm] <- summary.aov(aov(eff.perm.H0int ~ data$Smoking.status + exercise + data$Smoking.status:exercise))[[1]][3,4]
}
sum(T_int >= T0_int)/B

#test two factors
B <- 1000
T0_exerc <- summary.aov(aov(data$Sleep.efficiency ~ data$Smoking.status + exercise))[[1]][2,4]
aov.H0exerc <- aov(data$Sleep.efficiency ~ data$Smoking.status)
residuals.H0exerc <- aov.H0exerc$residuals

T0_smoke <- summary.aov(aov(data$Sleep.efficiency ~ data$Smoking.status + exercise))[[1]][1,4]
aov.H0smoke <- aov(data$Sleep.efficiency ~ exercise)
residuals.H0smoke <- aov.H0smoke$residuals

T_exerc <- T_smoke <- numeric(B)
for(perm in 1:B){
permutation <- sample(n)
# Test exerc
effic.perm.H0exerc <- aov.H0exerc$fitted + residuals.H0exerc[permutation]
T_exerc[perm] <- summary.aov(aov(effic.perm.H0exerc ~ data$Smoking.status + exercise))[[1]][2,4]
# Test smoke
effic.perm.H0smoke <- aov.H0smoke$fitted + residuals.H0smoke[permutation]
T_smoke[perm] <- summary.aov(aov(effic.perm.H0smoke ~ data$Smoking.status + exercise))[[1]][1,4]
}
sum(T_exerc >= T0_exerc)/B
sum(T_smoke >= T0_smoke)/B
```






