---
output: html_document
editor_options: 
  chunk_output_type: console
---

We have a "weird" cluster of people with a VERY high wage and we observe
some kind of reverse bathtub behavior. What happens if I try to model it
linearly?

```{r}
datajitt <- data %>%
  mutate(across(c(6, 11, 12, 13, 14, 15), ~  pmax(0,
                                          jitter(as.numeric(.), length(levels(as.factor(.))),0)
                                          )
                ))

# datajitt <- data %>%
#   mutate(across(c(6, 12, 13, 15), ~  pmax(0,
#                                           jitter(as.numeric(.), length(levels(as.factor(.))))
#                                           )
#                 ))


plot(datajitt)
```

# variabili su cui sono interessato svolgere le analisi
```{r}
plot(datajitt$Sleep.efficiency, datajitt$Deep.sleep.percentage) 
plot(datajitt$Age, datajitt$Exercise.frequency)
plot(datajitt$Awakenings, datajitt$Exercise.frequency)
plot(datajitt$Exercise.frequency, datajitt$Sleep.efficiency)
```

# Global Estimation
## Regressione polinomiale
```{r}
plot(datajitt$Sleep.efficiency, datajitt$Deep.sleep.percentage) #prova con un taglio

#polynomial reg
m_list_eff_deep = lapply(1:10, function(degree){
   with(datajitt,lm(Sleep.efficiency ~ poly(Deep.sleep.percentage,degree=degree)))})
do.call(anova, m_list_eff_deep)


m_list_eff_deep_raw <- lapply(1:10, function(degree) {
  with(datajitt, lm(Sleep.efficiency ~ poly(Deep.sleep.percentage, degree = degree, raw = TRUE)))
})

# summary(m_list_eff_deep[[8]]) # non tutte le variabili significative
# summary(m_list_eff_deep[[7]]) # non tutte le variabili significative
# summary(m_list_eff_deep[[6]]) # non tutte le variabili significative
# summary(m_list_eff_deep[[5]]) # Adjusted R-squared:  0.7076
# summary(m_list_eff_deep[[4]]) # Adjusted R-squared:  0.6951


# summary(m_list_eff_deep_raw[[8]]) # tutte le variabili significative pvalue <0.02
# summary(m_list_eff_deep_raw[[7]])
# summary(m_list_eff_deep_raw[[6]])
summary(m_list_eff_deep_raw[[5]]) # tutte le variabili significative pvalue <0.02
summary(m_list_eff_deep_raw[[4]])

#Scegliamo grado 8
par(mfrow=c(2,2))
plot(m_list_eff_deep[[5]])
par(mfrow=c(1,1))

deep_sleep_grid <- seq(min(datajitt$Deep.sleep.percentage), max(datajitt$Deep.sleep.percentage),
                       length.out = 100)

preds <- predict(m_list_eff_deep_raw[[5]], list(Deep.sleep.percentage = deep_sleep_grid), se.fit = TRUE)
se_bands <- cbind(preds$fit + 2 * preds$se.fit, preds$fit - 2 * preds$se.fit)

plot(datajitt$Deep.sleep.percentage, datajitt$Sleep.efficiency, 
     xlim = range(deep_sleep_grid), cex = 0.5, col = "darkgrey", 
     main = "Polynomial Regression - Degree 5", 
     xlab = "Deep Sleep Percentage", ylab = "Sleep Efficiency")

lines(deep_sleep_grid, preds$fit, lwd = 2, col = "blue")
matlines(deep_sleep_grid, se_bands, lwd = 1, col = "blue", lty = 3)
```

```{r}
plot(datajitt$Age, datajitt$Exercise.frequency)

m_list_age_exer <- lapply(1:8, function(degree){
   with(datajitt, lm(Exercise.frequency ~ poly(Age, degree = degree)))
})
do.call(anova, m_list_age_exer)

m_list_age_exer_raw <- lapply(1:8, function(degree) {
  with(datajitt, lm(Exercise.frequency ~ poly(Age+0, degree = degree, raw = TRUE)))
})
# 
# summary(m_list_age_exer[[6]])
# summary(m_list_age_exer[[5]])
# summary(m_list_age_exer[[4]])
# summary(m_list_age_exer[[3]])
# summary(m_list_age_exer[[2]])
 
# summary(m_list_age_exer_raw[[6]])
# summary(m_list_age_exer_raw[[5]]) # tutte significative, good
# summary(m_list_age_exer_raw[[4]])
summary(m_list_age_exer_raw[[3]])
summary(m_list_age_exer_raw[[2]])

par(mfrow=c(2,2))
plot(m_list_age_exer[[2]])
par(mfrow=c(1,1))

library(car)
influencePlot(m_list_age_exer[[2]])
id_out <- as.numeric(rownames(influencePlot(m_list_age_exer[[2]], fill = F)))

m_list_age_exer_raw <- lapply(1:8, function(degree) {
  with(datajitt[-id_out,], lm(Exercise.frequency ~ poly(Age, degree = degree, raw = TRUE)))
})
do.call(anova, m_list_age_exer_raw)

# summary(m_list_age_exer_raw[[6]])
# summary(m_list_age_exer_raw[[5]])
# summary(m_list_age_exer_raw[[4]])
summary(m_list_age_exer_raw[[3]])
summary(m_list_age_exer_raw[[2]])


age_grid <- seq(min(datajitt$Age), max(datajitt$Age), length.out = 100)
preds_age_exer <- predict(m_list_age_exer_raw[[2]], list(Age = age_grid), se.fit = TRUE)
se_bands_age_exer <- cbind(preds_age_exer$fit + 2 * preds_age_exer$se.fit, preds_age_exer$fit - 2 * preds_age_exer$se.fit)

plot(datajitt$Age, datajitt$Exercise.frequency, 
     xlim = range(age_grid), cex = 0.5, col = "darkgrey", 
     main = "Polynomial Regression - Degree 2", 
     xlab = "Age", ylab = "Exercise Frequency")
lines(age_grid, preds_age_exer$fit, lwd = 2, col = "blue")
matlines(age_grid, se_bands_age_exer, lwd = 1, col = "blue", lty = 3)
```

```{r}
plot(datajitt$Awakenings, datajitt$Exercise.frequency)

m_list_exer_awak <- lapply(1:8, function(degree){
   with(datajitt, lm(Awakenings ~ poly(Exercise.frequency, degree = degree)))
})
do.call(anova, m_list_exer_awak) #pvalue 0.09587 aggiungendo una covariata da 1 -> 2; non ci sono motivi per aggiungerla

m_list_exer_awak_raw <- lapply(1:8, function(degree) {
  with(datajitt, lm(Awakenings ~ poly(Exercise.frequency, degree = degree, raw = TRUE)))
})

# summary(m_list_exer_awak[[1]])
# summary(m_list_exer_awak[[2]])
# summary(m_list_exer_awak[[3]])

summary(m_list_exer_awak_raw[[1]]) #ALL SIGNIFICATIVE 
summary(m_list_exer_awak_raw[[2]]) # ^2 pvalue >0.05
summary(m_list_exer_awak_raw[[3]]) # nessuna significativa

par(mfrow=c(2,2))
plot(m_list_exer_awak[[2]])
par(mfrow=c(1,1))

exer_grid <- seq(min(datajitt$Exercise.frequency), max(datajitt$Exercise.frequency), length.out = 300)
preds_exer_awak <- predict(m_list_exer_awak[[1]], list(Exercise.frequency = exer_grid), se.fit = TRUE)
se_bands_exer_awak <- cbind(preds_exer_awak$fit + 2 * preds_exer_awak$se.fit, preds_exer_awak$fit - 2 * preds_exer_awak$se.fit)

plot(datajitt$Exercise.frequency, datajitt$Awakenings, 
     xlim = range(exer_grid), cex = 0.5, col = "darkgrey", 
     main = "Degree 1 - Linear Regression", 
     xlab = "Exercise Frequency", ylab = "Awakenings")
lines(exer_grid, preds_exer_awak$fit, lwd = 2, col = "blue")
matlines(exer_grid, se_bands_exer_awak, lwd = 1, col = "blue", lty = 3)


plot(datajitt$Exercise.frequency, datajitt$Sleep.efficiency)

m_list_exer_sleep <- lapply(1:8, function(degree){
   with(datajitt, lm(Sleep.efficiency ~ poly(Exercise.frequency, degree = degree)))
})
do.call(anova, m_list_exer_sleep) #pvalue tutti alti utilizzando anova test, non ha senso aggiungere covariate

m_list_exer_sleep_raw <- lapply(1:8, function(degree) {
  with(datajitt, lm(Sleep.efficiency ~ poly(Exercise.frequency, degree = degree, raw = TRUE)))
})

# summary(m_list_exer_sleep[[6]])
# summary(m_list_exer_sleep[[5]])
# summary(m_list_exer_sleep[[4]])
# summary(m_list_exer_sleep[[3]])
# summary(m_list_exer_sleep[[2]])
# 
# summary(m_list_exer_sleep_raw[[6]])
# summary(m_list_exer_sleep_raw[[5]])
# summary(m_list_exer_sleep_raw[[4]])
# summary(m_list_exer_sleep_raw[[3]])
# summary(m_list_exer_sleep_raw[[2]])

par(mfrow=c(2,2))
plot(m_list_exer_sleep[[1]])
par(mfrow=c(1,1))

summary(m_list_exer_sleep[[1]])

exer_grid <- seq(min(datajitt$Exercise.frequency), max(datajitt$Exercise.frequency), length.out = 100)
preds_exer_sleep <- predict(m_list_exer_sleep[[1]], list(Exercise.frequency = exer_grid), se.fit = TRUE)
se_bands_exer_sleep <- cbind(preds_exer_sleep$fit + 2 * preds_exer_sleep$se.fit, preds_exer_sleep$fit - 2 * preds_exer_sleep$se.fit)

plot(datajitt$Exercise.frequency, datajitt$Sleep.efficiency, 
     xlim = range(exer_grid), cex = 0.5, col = "darkgrey", 
     main = "Degree 1 - Linear Regression", 
     xlab = "Exercise Frequency", ylab = "Sleep Efficiency")
lines(exer_grid, preds_exer_sleep$fit, lwd = 2, col = "blue")
matlines(exer_grid, se_bands_exer_sleep, lwd = 1, col = "blue", lty = 3)

```

## Regressione Logistica carina
```{r}
m_list_logit_eff = lapply(1:6,function(degree){with(datajitt, glm(I(Sleep.efficiency>0.8) ~ poly(Deep.sleep.percentage,degree=degree),family='binomial'))})
do.call(what = anova, c(list(test="Chisq"), m_list_logit_eff))

deep_sleep_grid <- seq(min(datajitt$Deep.sleep.percentage), max(datajitt$Deep.sleep.percentage), length.out = 100)
preds <- predict(m_list_logit_eff[[2]], list(Deep.sleep.percentage = deep_sleep_grid), se.fit = TRUE)

logistic <- function(x) exp(x) / (1 + exp(x))
pfit <- sapply(preds$fit, FUN = logistic)

se_bands_logit <- cbind(preds$fit + 1.5 * preds$se.fit, preds$fit - 1.5 * preds$se.fit)
se_bands <- logistic(se_bands_logit) 

# plot(datajitt$Deep.sleep.percentage, I(datajitt$Sleep.efficiency > 0.8), 
#      xlim = range(deep_sleep_grid), type = "n", ylim = c(0, 1), 
#      xlab = "Deep Sleep Percentage", ylab = "P(Sleep Efficiency > 0.8)", 
#      main = "Grado 4 - Modello Logistico")
# 
# points(jitter(datajitt$Deep.sleep.percentage), (datajitt$Sleep.efficiency > 0.8), 
#        cex = 1, pch = 'I', col = "black")
# 
# lines(deep_sleep_grid, pfit, lwd = 2, col = "blue4")
# matlines(deep_sleep_grid, se_bands, lwd = 1, col = "blue4", lty = 3)


summary(m_list_logit_eff[[2]])
preds2 <- predict(m_list_logit_eff[[2]], list(Deep.sleep.percentage = deep_sleep_grid), se.fit = TRUE)

pfit2 <- sapply(preds2$fit, FUN = logistic)
se_bands_logit2 <- cbind(preds2$fit + 0.5 * preds2$se.fit, preds2$fit - 0.5 * preds2$se.fit)
se_bands2 <- logistic(se_bands_logit2)

plot(datajitt$Deep.sleep.percentage, I(datajitt$Sleep.efficiency > 0.8), 
     xlim = range(deep_sleep_grid), type = "n", ylim = c(0, 1), 
     xlab = "Deep Sleep Percentage", ylab = "P(Sleep Efficiency > 0.8)", 
     main = "Logistic Model - Degree 2")
points(jitter(datajitt$Deep.sleep.percentage), (datajitt$Sleep.efficiency > 0.8), 
       cex = 1, pch = 'I', col = "black")
lines(deep_sleep_grid, pfit2, lwd = 2, col = "aquamarine4")
matlines(deep_sleep_grid, se_bands2, lwd = 1, col = "aquamarine4", lty = 3)
```


```{r}
# datajitt$Gender <- as.factor(datajitt$Gender)
# 
# m_list_logit_eff <- with(datajitt, 
#        glm(I(Sleep.efficiency > 0.8) ~ poly(as.numeric(Gender), degree = 1), 
#            family = "binomial"))
# 
# summary(m_list_logit_eff)
# 
# deep_sleep_grid <- seq(min(datajitt$Deep.sleep.percentage), max(datajitt$Deep.sleep.percentage), length.out = 100)
# 
# # Predict Probabilities for the Best Model (Degree 4)
# preds <- predict(m_list_logit_eff, newdata = data.frame(
#   Deep.sleep.percentage = deep_sleep_grid,
#   Gender = as.factor(rep("Male", length(deep_sleep_grid)))  # Set Gender to "Male" for prediction
# ), type = "link", se.fit = TRUE)
# 
# # Logistic Function to Transform Log-Odds to Probabilities
# logistic <- function(x) exp(x) / (1 + exp(x))
# 
# # Calculate Fitted Probabilities and Confidence Bands
# pfit <- sapply(preds$fit, FUN = logistic)
# se_bands_logit <- cbind(preds$fit + 1.5 * preds$se.fit, preds$fit - 1.5 * preds$se.fit)
# se_bands <- logistic(se_bands_logit)
# 
# # Plot for Degree 4 Model
# plot(datajitt$Deep.sleep.percentage, I(datajitt$Sleep.efficiency > 0.8), 
#      xlim = range(deep_sleep_grid), type = "n", ylim = c(0, 1), 
#      xlab = "Deep Sleep Percentage", ylab = "P(Sleep Efficiency > 0.8)", 
#      main = "Grado 4 - Modello Logistico")
# points(jitter(datajitt$Deep.sleep.percentage), (datajitt$Sleep.efficiency > 0.8), 
#        cex = 1, pch = 'I', col = "black")
# lines(deep_sleep_grid, pfit, lwd = 2, col = "blue4")
# matlines(deep_sleep_grid, se_bands, lwd = 1, col = "blue4", lty = 3)
```

```{r}
m_list_logit_ag <- lapply(1:6,function(degree){with(datajitt %>% filter(Age>20), glm(I(Awakenings>1.5) ~ poly(Age,degree=degree),family='binomial'))})
summary(m_list_logit_ag[[1]])

ag_grid <- seq(min(datajitt$Age), max(datajitt$Age), length.out = 300)
preds <- predict(m_list_logit_ag[[1]], list(Age = ag_grid), se.fit = TRUE)

logistic <- function(x) exp(x) / (1 + exp(x))
pfit <- sapply(preds$fit, FUN = logistic)

se_bands_logit <- cbind(preds$fit + 1.5 * preds$se.fit, preds$fit - 1.5 * preds$se.fit)
se_bands <- logistic(se_bands_logit) 

plot(datajitt$Age, I(datajitt$Age > 0.8),
     xlim = range(ag_grid), type = "n", ylim = c(0, 1),
     xlab = "Age", ylab = "P(Awakenings > 1.5)",
     main = "Logistic Model")

points(jitter(datajitt$Age), (datajitt$Awakenings > 1.5),
       cex = 1, pch = 'I', col = "black")

lines(ag_grid, pfit, lwd = 2, col = "violetred4")
matlines(ag_grid, se_bands, lwd = 1, col = "violetred4", lty = 3)
```


```{r}
m_list_logit_exer <- lapply(1:6,function(degree){with(datajitt, glm(I(Sleep.efficiency>0.8) ~ poly(Exercise.frequency,degree=degree),family='binomial'))})
summary(m_list_logit_exer[[1]])

exer_grid <- seq(min(datajitt$Exercise.frequency), max(datajitt$Exercise.frequency), length.out = 300)
preds <- predict(m_list_logit_exer[[1]], list(Exercise.frequency = exer_grid), se.fit = TRUE)

logistic <- function(x) exp(x) / (1 + exp(x))
pfit <- sapply(preds$fit, FUN = logistic)

se_bands_logit <- cbind(preds$fit + 1.5 * preds$se.fit, preds$fit - 1.5 * preds$se.fit)
se_bands <- logistic(se_bands_logit) 

plot(datajitt$Exercise.frequency, I(datajitt$Sleep.efficiency > 0.8),
     xlim = range(exer_grid), type = "n", ylim = c(0, 1),
     xlab = "Exercise frequency", ylab = "P(Sleep Efficiency > 0.8)",
     main = "Logistic Model")

points(jitter(datajitt$Exercise.frequency), (datajitt$Sleep.efficiency > 0.8),
       cex = 1, pch = 'I', col = "black")

lines(exer_grid, pfit, lwd = 2, col = "chocolate")
matlines(exer_grid, se_bands, lwd = 1, col = "chocolate", lty = 3)
```

## Step function:

```{r}
### plot(datajitt$Sleep.efficiency, datajitt$Deep.sleep.percentage)

m_cut_eff_deep <- with(datajitt, lm(Sleep.efficiency ~ cut(Deep.sleep.percentage,breaks = c(-Inf,45,+Inf))))
broom::tidy(summary(m_cut_eff_deep)) %>% 
   dplyr::mutate(term=ifelse(term=="(Intercept)","(-Inf,45]", "(45, Inf]"))

deep_sleep_grid <- seq(min(datajitt$Deep.sleep.percentage), max(datajitt$Deep.sleep.percentage), by = 0.3)
preds <- predict(m_cut_eff_deep, list(Deep.sleep.percentage = deep_sleep_grid), se.fit = TRUE)
se_bands <- cbind(preds$fit + 2 * preds$se.fit, preds$fit - 2 * preds$se.fit)

with(datajitt, 
     plot(Deep.sleep.percentage, Sleep.efficiency, cex = 0.5, main = "Custom Cut Fit", 
          xlab = "Deep Sleep Percentage", ylab = "Sleep Efficiency"))

lines(deep_sleep_grid, preds$fit, lwd = 2, col = "blue")
matlines(deep_sleep_grid, se_bands, lwd = 1, col = "blue", lty = 3)

### plot(datajitt$Age, datajitt$Exercise.frequency)
m_cut_age_exer <- with(datajitt, 
                       lm(Exercise.frequency ~ cut(Age, breaks = 8)))
summary(m_cut_age_exer)

age_grid <- seq(min(datajitt$Age), max(datajitt$Age), by = 0.01)

preds_age_exer <- predict(m_cut_age_exer, list(Age = age_grid), se.fit = TRUE)
se_bands_age_exer <- cbind(preds_age_exer$fit + 2 * preds_age_exer$se.fit, preds_age_exer$fit - 2 * preds_age_exer$se.fit)

with(datajitt, 
     plot(Age, Exercise.frequency, cex = 0.5, main = "Custom Cut Fit (Age vs Exercise Frequency)", 
          xlab = "Age", ylab = "Exercise Frequency"))
lines(age_grid, preds_age_exer$fit, lwd = 2, col = "blue")
matlines(age_grid, se_bands_age_exer, lwd = 1, col = "blue", lty = 3)

### plot(datajitt$Awakenings, datajitt$Exercise.frequency)
m_cut_awak_exer <- with(datajitt, 
                        lm(Exercise.frequency ~ cut(Awakenings, breaks = c(-Inf, 1.5, +Inf))))
summary(m_cut_awak_exer)

awak_grid <- seq(min(datajitt$Awakenings), max(datajitt$Awakenings), by = 0.01)

preds_awak_exer <- predict(m_cut_awak_exer, list(Awakenings = awak_grid), se.fit = TRUE)
se_bands_awak_exer <- cbind(preds_awak_exer$fit + 2 * preds_awak_exer$se.fit, preds_awak_exer$fit - 2 * preds_awak_exer$se.fit)

with(datajitt, 
     plot(Awakenings, Exercise.frequency, cex = 0.5, main = "Custom Cut Fit (Awakenings vs Exercise Frequency)", 
          xlab = "Awakenings", ylab = "Exercise Frequency"))
lines(awak_grid, preds_awak_exer$fit, lwd = 2, col = "blue")
matlines(awak_grid, se_bands_awak_exer, lwd = 1, col = "blue", lty = 3)

### plot(datajitt$Exercise.frequency, datajitt$Sleep.efficiency)
m_cut_exer_sleep <- with(datajitt, 
                         lm(Sleep.efficiency ~ cut(Exercise.frequency, breaks = c(-Inf, 2, 4, +Inf))))
summary(m_cut_exer_sleep)

exer_grid <- seq(min(datajitt$Exercise.frequency), max(datajitt$Exercise.frequency), by = 0.01)

preds_exer_sleep <- predict(m_cut_exer_sleep, list(Exercise.frequency = exer_grid), se.fit = TRUE)
se_bands_exer_sleep <- cbind(preds_exer_sleep$fit + 2 * preds_exer_sleep$se.fit, preds_exer_sleep$fit - 2 * preds_exer_sleep$se.fit)

with(datajitt, 
     plot(Exercise.frequency, Sleep.efficiency, cex = 0.5, main = "Custom Cut Fit (Exercise Frequency vs Sleep Efficiency)", 
          xlab = "Exercise Frequency", ylab = "Sleep Efficiency"))
lines(exer_grid, preds_exer_sleep$fit, lwd = 2, col = "blue")
matlines(exer_grid, se_bands_exer_sleep, lwd = 1, col = "blue", lty = 3)
```

# Local Regression
```{r message=FALSE, warning=FALSE}
library(ISLR2)
library(car)
library(np)
library(splines)
library(fda)
library(magrittr)
library(KernSmooth)
```

## Uniform Kernel
```{r}
## plot(datajitt$Deep.sleep.percentage, datajitt$Sleep.efficiency)
m_loc_deep_eff_25 <- with(datajitt, 
                          npreg(Sleep.efficiency ~ Deep.sleep.percentage,
                                ckertype = 'uniform',
                                bws = 25)
)

m_loc_deep_eff_15 <- with(datajitt, 
                          npreg(Sleep.efficiency ~ Deep.sleep.percentage,
                                ckertype = 'uniform',
                                bws = 15)
)

m_loc_deep_eff_8.5 <- with(datajitt, 
                           npreg(Sleep.efficiency ~ Deep.sleep.percentage,
                                 ckertype = 'uniform',
                                 bws = 8.5)
)

deep_newdata <- data.frame(Deep.sleep.percentage = with(datajitt, 
                                                       seq(range(Deep.sleep.percentage)[1], 
                                                          range(Deep.sleep.percentage)[2],
                                                          by = 0.005))
)


summary(m_loc_deep_eff_25)
summary(m_loc_deep_eff_15)
summary(m_loc_deep_eff_8.5)

preds_25 <- predict(m_loc_deep_eff_25, newdata = deep_newdata, se = TRUE)
preds_15 <- predict(m_loc_deep_eff_15, newdata = deep_newdata, se = TRUE)
preds_8.5 <- predict(m_loc_deep_eff_8.5, newdata = deep_newdata, se = TRUE)

se_bands_25 <- cbind(preds_25$fit + 2 * preds_25$se.fit, preds_25$fit - 2 * preds_25$se.fit)
se_bands_15 <- cbind(preds_15$fit + 2 * preds_15$se.fit, preds_15$fit - 2 * preds_15$se.fit)
se_bands_8.5 <- cbind(preds_8.5$fit + 2 * preds_8.5$se.fit, preds_8.5$fit - 2 * preds_8.5$se.fit)

with(datajitt,
     plot(Deep.sleep.percentage, Sleep.efficiency, 
          cex = 0.5,
          col = "gray35",
          main = "Local Averaging - Uniform Kernel",
          xlab = "Deep Sleep Percentage",
          ylab = "Sleep Efficiency"
     )
)

lines(deep_newdata$Deep.sleep.percentage, preds_25$fit, lwd = 2, col = "blue4", lty = 1)
matlines(deep_newdata$Deep.sleep.percentage, se_bands_25, lwd = 1, col = "blue4", lty = 3)

lines(deep_newdata$Deep.sleep.percentage, preds_15$fit, lwd = 2, col = "red4", lty = 1)
matlines(deep_newdata$Deep.sleep.percentage, se_bands_15, lwd = 1, col = "red4", lty = 3)

lines(deep_newdata$Deep.sleep.percentage, preds_8.5$fit, lwd = 2, col = "darkgreen", lty = 1)
matlines(deep_newdata$Deep.sleep.percentage, se_bands_8.5, lwd = 1, col = "darkgreen", lty = 3)

legend("topleft", 
       legend = c("bws = 25", "bws = 15", "bws = 8.5"), 
       col = c("blue4", "red4", "green4"), 
       lty = c(1, 1, 1), 
       lwd = 2, 
       title = "Bandwidths",
       bty = "n")

```

## Uniform Kernel

```{r}
## plot(datajitt$Age, datajitt$Exercise.frequency)
m_loc_age_exer_7 <- with(datajitt, 
                         npreg(Exercise.frequency ~ Age,
                               ckertype = 'uniform',
                               bws = 12)
)

m_loc_age_exer_4.1 <- with(datajitt, 
                           npreg(Exercise.frequency ~ Age,
                                 ckertype = 'uniform',
                                 bws = 6)
)

m_loc_age_exer_2.5 <- with(datajitt, 
                           npreg(Exercise.frequency ~ Age,
                                 ckertype = 'uniform',
                                 bws = 2.3)
)

age_newdata <- data.frame(Age = with(datajitt, 
                                     seq(range(Age)[1], range(Age)[2], by = 0.1))
)

preds_age_7 <- predict(m_loc_age_exer_7, newdata = age_newdata, se = TRUE)
preds_age_4.1 <- predict(m_loc_age_exer_4.1, newdata = age_newdata, se = TRUE)
preds_age_2.5 <- predict(m_loc_age_exer_2.5, newdata = age_newdata, se = TRUE)

se_bands_age_7 <- cbind(preds_age_7$fit + 2 * preds_age_7$se.fit, preds_age_7$fit - 2 * preds_age_7$se.fit)
se_bands_age_4.1 <- cbind(preds_age_4.1$fit + 2 * preds_age_4.1$se.fit, preds_age_4.1$fit - 2 * preds_age_4.1$se.fit)
se_bands_age_2.5 <- cbind(preds_age_2.5$fit + 2 * preds_age_2.5$se.fit, preds_age_2.5$fit - 2 * preds_age_2.5$se.fit)

with(datajitt,
     plot(Age, Exercise.frequency, 
          cex = 0.5,
          col = "gray35",
          main = "Local Averaging - Age vs Exercise Frequency",
          xlab = "Age",
          ylab = "Exercise Frequency"
     )
)


lines(age_newdata$Age, preds_age_7$fit, lwd = 2, col = "blue4", lty = 1)
matlines(age_newdata$Age, se_bands_age_7, lwd = 1, col = "blue4", lty = 3)
lines(age_newdata$Age, preds_age_4.1$fit, lwd = 2, col = "red4", lty = 1)
matlines(age_newdata$Age, se_bands_age_4.1, lwd = 1, col = "red4", lty = 3)
lines(age_newdata$Age, preds_age_2.5$fit, lwd = 2, col = "green4", lty = 1)
matlines(age_newdata$Age, se_bands_age_2.5, lwd = 1, col = "green4", lty = 3)

legend("topleft", 
       legend = c("bws = 12", "bws = 6", "bws = 2.3"), 
       col = c("blue4", "red4", "green4"), 
       lty = c(1, 1, 1), 
       lwd = 2, 
       title = "Bandwidths",
       bty = "n"
)
## plot(datajitt$Awakenings, datajitt$Exercise.frequency)
m_loc_awak_exer_1.9 <- with(datajitt, 
                            npreg(Exercise.frequency ~ Awakenings,
                                  ckertype = 'uniform',
                                  bws = 1.9)
)

m_loc_awak_exer_0.4 <- with(datajitt, 
                            npreg(Exercise.frequency ~ Awakenings,
                                  ckertype = 'uniform',
                                  bws = 0.4)
)

awak_newdata <- data.frame(Awakenings = with(datajitt, 
                                            seq(range(Awakenings)[1], range(Awakenings)[2], by = 0.01))
)

preds_awak_1.9 <- predict(m_loc_awak_exer_1.9, newdata = awak_newdata, se = TRUE)
preds_awak_0.4 <- predict(m_loc_awak_exer_0.4, newdata = awak_newdata, se = TRUE)

se_bands_awak_1.9 <- cbind(preds_awak_1.9$fit + 2 * preds_awak_1.9$se.fit, preds_awak_1.9$fit - 2 * preds_awak_1.9$se.fit)
se_bands_awak_0.4 <- cbind(preds_awak_0.4$fit + 2 * preds_awak_0.4$se.fit, preds_awak_0.4$fit - 2 * preds_awak_0.4$se.fit)

with(datajitt,
     plot(Awakenings, Exercise.frequency, 
          cex = 0.5,
          col = "gray35",
          main = "Local Averaging - Awakenings vs Exercise Frequency",
          xlab = "Awakenings",
          ylab = "Exercise Frequency"
     )
)

lines(awak_newdata$Awakenings, preds_awak_1.9$fit, lwd = 2, col = "blue4", lty = 1)
matlines(awak_newdata$Awakenings, se_bands_awak_1.9, lwd = 1, col = "blue4", lty = 3)

lines(awak_newdata$Awakenings, preds_awak_0.4$fit, lwd = 2, col = "green4", lty = 1)
matlines(awak_newdata$Awakenings, se_bands_awak_0.4, lwd = 1, col = "green4", lty = 3)

legend("topleft", 
       legend = c("bws = 1.9", "bws = 0.4"), 
       col = c("blue4", "green4"), 
       lty = c(1, 1), 
       lwd = 2, 
       title = "Bandwidths",
       bty = "n"
)

## plot(datajitt$Exercise.frequency, datajitt$Sleep.efficiency)
m_loc_exer_sleep_2.2 <- with(datajitt,
                             npreg(Sleep.efficiency ~ Exercise.frequency,
                                   ckertype = 'uniform',
                                   bws = 2.2)
)

m_loc_exer_sleep_0.8 <- with(datajitt, 
                             npreg(Sleep.efficiency ~ Exercise.frequency,
                                   ckertype = 'uniform',
                                   bws = 0.8)
)

m_loc_exer_sleep_0.3 <- with(datajitt, 
                             npreg(Sleep.efficiency ~ Exercise.frequency,
                                   ckertype = 'uniform',
                                   bws = 0.3)
)

exer_newdata <- data.frame(Exercise.frequency = with(datajitt, 
                                                    seq(range(Exercise.frequency)[1], range(Exercise.frequency)[2], by = 0.01))
)

preds_exer_2.2 <- predict(m_loc_exer_sleep_2.2, newdata = exer_newdata, se = TRUE)
preds_exer_0.8 <- predict(m_loc_exer_sleep_0.8, newdata = exer_newdata, se = TRUE)
preds_exer_0.3 <- predict(m_loc_exer_sleep_0.3, newdata = exer_newdata, se = TRUE)

se_bands_exer_2.2 <- cbind(preds_exer_2.2$fit + 2 * preds_exer_2.2$se.fit, preds_exer_2.2$fit - 2 * preds_exer_2.2$se.fit)
se_bands_exer_0.8 <- cbind(preds_exer_0.8$fit + 2 * preds_exer_0.8$se.fit, preds_exer_0.8$fit - 2 * preds_exer_0.8$se.fit)
se_bands_exer_0.3 <- cbind(preds_exer_0.3$fit + 2 * preds_exer_0.3$se.fit, preds_exer_0.3$fit - 2 * preds_exer_0.3$se.fit)

with(datajitt,
     plot(Exercise.frequency, Sleep.efficiency, 
          cex = 0.5,
          col = "gray35",
          main = "Local Averaging - Exercise Frequency vs Sleep Efficiency",
          xlab = "Exercise Frequency",
          ylab = "Sleep Efficiency"
     )
)

lines(exer_newdata$Exercise.frequency, preds_exer_2.2$fit, lwd = 2, col = "blue4", lty = 1)
matlines(exer_newdata$Exercise.frequency, se_bands_exer_2.2, lwd = 1, col = "blue4", lty = 3)
lines(exer_newdata$Exercise.frequency, preds_exer_0.8$fit, lwd = 2, col = "red4", lty = 1)
matlines(exer_newdata$Exercise.frequency, se_bands_exer_0.8, lwd = 1, col = "red4", lty = 3)
lines(exer_newdata$Exercise.frequency, preds_exer_0.3$fit, lwd = 2, col = "green4", lty = 1)
matlines(exer_newdata$Exercise.frequency, se_bands_exer_0.3, lwd = 1, col = "green4", lty = 3)

legend("topleft", 
       legend = c("bws = 2.2", "bws = 0.8", "bws = 0.3"), 
       col = c("blue4", "red4", "green4"), 
       lty = c(1, 1, 1), 
       lwd = 2, 
       title = "Bandwidths",
       bty = "n"
)
```

## Kernel gaussian

```{r}
## plot(datajitt$Sleep.efficiency, datajitt$Deep.sleep.percentage)
m_loc_deep_sleep_gaus <- with(datajitt, 
                              npreg(Sleep.efficiency ~ Deep.sleep.percentage,
                                    ckertype = 'gaussian',
                                    bws = 5.8)
)
deep_newdata <- data.frame(Deep.sleep.percentage = with(datajitt, 
                                                       seq(range(Deep.sleep.percentage)[1], range(Deep.sleep.percentage)[2], by = 0.1))
)
preds_deep_gaus <- predict(m_loc_deep_sleep_gaus, newdata = deep_newdata, se = TRUE)
se_bands_deep_gaus <- cbind(preds_deep_gaus$fit + 2 * preds_deep_gaus$se.fit, preds_deep_gaus$fit - 2 * preds_deep_gaus$se.fit)

m_loess_deep <- loess(Sleep.efficiency ~ Deep.sleep.percentage, # Modello LOESS
                      span = 0.35, 
                      degree = 0,  # Estimatore di Nadaraya-Watson
                      data = datajitt)
summary(m_loc_deep_sleep_gaus)
summary(m_loess_deep)

preds_loess_deep <- predict(m_loess_deep, newdata = deep_newdata, se = TRUE)
se_bands_loess_deep <- cbind(preds_loess_deep$fit + 2 * preds_loess_deep$se.fit, preds_loess_deep$fit - 2 * preds_loess_deep$se.fit)

# Plot
with(datajitt,
     plot(Deep.sleep.percentage, Sleep.efficiency, 
          cex = 0.5,
          col = "gray35",
          main = "Deep Sleep Percentage vs Sleep Efficiency",
          xlab = "Deep Sleep Percentage",
          ylab = "Sleep Efficiency"
     )
)
lines(deep_newdata$Deep.sleep.percentage, preds_deep_gaus$fit, lwd = 2, col = "blue4", lty = 1)
matlines(deep_newdata$Deep.sleep.percentage, se_bands_deep_gaus, lwd = 1, col = "blue4", lty = 3)
lines(deep_newdata$Deep.sleep.percentage, preds_loess_deep$fit, lwd = 2, col = "orangered", lty = 1)
matlines(deep_newdata$Deep.sleep.percentage, se_bands_loess_deep, lwd = 1, col = "orangered", lty = 3)
legend("topleft", 
       legend = c("Gaussian Kernel", "LOESS"), 
       col = c("blue4", "orangered"), 
       lty = c(1, 1), 
       lwd = 2, 
       title = "Methods",
       bty = "n"
)

## plot(datajitt$Age, datajitt$Exercise.frequency)
m_loc_age_exer_gaus <- with(datajitt, 
                            npreg(Exercise.frequency ~ Age,
                                  ckertype = 'gaussian',
                                  bws = 6)
)
age_newdata <- data.frame(Age = with(datajitt, 
                                     seq(range(Age)[1], range(Age)[2], by = 0.1))
)
preds_age_gaus <- predict(m_loc_age_exer_gaus, newdata = age_newdata, se = TRUE)
se_bands_age_gaus <- cbind(preds_age_gaus$fit + 2 * preds_age_gaus$se.fit, preds_age_gaus$fit - 2 * preds_age_gaus$se.fit)

m_loess_age <- loess(Exercise.frequency ~ Age,
                     span = 0.35, 
                     degree = 0,  # Estimatore di Nadaraya-Watson
                     data = datajitt)

summary(m_loc_age_exer_gaus)
summary(m_loess_age)


preds_loess_age <- predict(m_loess_age, newdata = age_newdata, se = TRUE)
se_bands_loess_age <- cbind(preds_loess_age$fit + 2 * preds_loess_age$se.fit, preds_loess_age$fit - 2 * preds_loess_age$se.fit)

with(datajitt,
     plot(Age, Exercise.frequency, 
          cex = 0.5,
          col = "gray35",
          main = "Age vs Exercise Frequency",
          xlab = "Age",
          ylab = "Exercise Frequency"
     )
)
lines(age_newdata$Age, preds_age_gaus$fit, lwd = 2, col = "darkorchid", lty = 1)
matlines(age_newdata$Age, se_bands_age_gaus, lwd = 1, col = "darkorchid", lty = 3)
lines(age_newdata$Age, preds_loess_age$fit, lwd = 2, col = "orangered", lty = 1)
matlines(age_newdata$Age, se_bands_loess_age, lwd = 1, col = "orangered", lty = 3)
legend("topleft", 
       legend = c("Gaussian Kernel", "LOESS"), 
       col = c("darkorchid", "orangered"), 
       lty = c(1, 1), 
       lwd = 2, 
       title = "Methods",
       bty = "n"
)

## plot(datajitt$Awakenings, datajitt$Exercise.frequency)
m_loc_awak_exer_gaus <- with(datajitt, 
                             npreg(Exercise.frequency ~ Awakenings,
                                   ckertype = 'gaussian',
                                   bws = 0.4)
)
awak_newdata <- data.frame(Awakenings = with(datajitt, 
                                            seq(range(Awakenings)[1], range(Awakenings)[2], by = 0.01))
)
preds_awak_gaus <- predict(m_loc_awak_exer_gaus, newdata = awak_newdata, se = TRUE)
se_bands_awak_gaus <- cbind(preds_awak_gaus$fit + 2 * preds_awak_gaus$se.fit, preds_awak_gaus$fit - 2 * preds_awak_gaus$se.fit)

m_loess_awak <- loess(Exercise.frequency ~ Awakenings,
                      span = 0.35, 
                      degree = 0,  # Estimatore di Nadaraya-Watson
                      data = datajitt)

summary(m_loc_awak_exer_gaus)
summary(m_loess_awak)


preds_loess_awak <- predict(m_loess_awak, newdata = awak_newdata, se = TRUE)
se_bands_loess_awak <- cbind(preds_loess_awak$fit + 2 * preds_loess_awak$se.fit, preds_loess_awak$fit - 2 * preds_loess_awak$se.fit)

with(datajitt,
     plot(Awakenings, Exercise.frequency, 
          cex = 0.5,
          col = "gray35",
          main = "Awakenings vs Exercise Frequency",
          xlab = "Awakenings",
          ylab = "Exercise Frequency"
     )
)
lines(awak_newdata$Awakenings, preds_awak_gaus$fit, lwd = 2, col = "green4", lty = 1)
matlines(awak_newdata$Awakenings, se_bands_awak_gaus, lwd = 1, col = "green4", lty = 3)
lines(awak_newdata$Awakenings, preds_loess_awak$fit, lwd = 2, col = "orangered", lty = 1)
matlines(awak_newdata$Awakenings, se_bands_loess_awak, lwd = 1, col = "orangered", lty = 3)
legend("topleft", 
       legend = c("Gaussian Kernel", "LOESS"), 
       col = c("green4", "orangered"), 
       lty = c(1, 1), 
       lwd = 2, 
       title = "Methods",
       bty = "n")

## plot(datajitt$Exercise.frequency, datajitt$Sleep.efficiency)
m_loc_exer_sleep_gaus <- with(datajitt, 
                              npreg(Sleep.efficiency ~ Exercise.frequency,
                                    ckertype = 'gaussian',
                                    bws = 0.6)
)
exer_newdata <- data.frame(Exercise.frequency = with(datajitt, 
                                                    seq(range(Exercise.frequency)[1], range(Exercise.frequency)[2], by = 0.01))
)
summary(m_loc_exer_sleep_gaus)
preds_exer_gaus <- predict(m_loc_exer_sleep_gaus, newdata = exer_newdata, se = TRUE)
se_bands_exer_gaus <- cbind(preds_exer_gaus$fit + 2 * preds_exer_gaus$se.fit, preds_exer_gaus$fit - 2 * preds_exer_gaus$se.fit)

m_loess_exer <- loess(Sleep.efficiency ~ Exercise.frequency,
                      span = 0.35, 
                      degree = 0,  # Estimatore di Nadaraya-Watson
                      data = datajitt)

summary(m_loc_exer_sleep_gaus)
summary(m_loess_exer)
summary(datajitt$Sleep.efficiency)

preds_loess_exer <- predict(m_loess_exer, newdata = exer_newdata, se = TRUE)
se_bands_loess_exer <- cbind(preds_loess_exer$fit + 2 * preds_loess_exer$se.fit, preds_loess_exer$fit - 2 * preds_loess_exer$se.fit)

with(datajitt,
     plot(Exercise.frequency, Sleep.efficiency, 
          cex = 0.5,
          col = "gray35",
          main = "Local Regression with Gaussian Kernel",
          xlab = "Exercise Frequency",
          ylab = "Sleep Efficiency"
     )
)
lines(exer_newdata$Exercise.frequency, preds_exer_gaus$fit, lwd = 2, col = "red4", lty = 1)
matlines(exer_newdata$Exercise.frequency, se_bands_exer_gaus, lwd = 1, col = "red4", lty = 3)
# lines(exer_newdata$Exercise.frequency, preds_loess_exer$fit, lwd = 2, col = "orangered", lty = 1)
# matlines(exer_newdata$Exercise.frequency, se_bands_loess_exer, lwd = 1, col = "orangered", lty = 3)
# legend("topleft", 
#        legend = c("Gaussian Kernel", "LOESS"), 
#        col = c("red4", "orangered"), 
#        lty = c(1, 1), 
#        lwd = 2, 
#        title = "Methods",
#        bty = "n"
# )
```

```{r}
## plot(datajitt$Alcohol.consumption, datajitt$Sleep.efficiency)
m_loc_alc_sleep_gaus <- with(datajitt, 
                              npreg(Sleep.efficiency ~ Alcohol.consumption,
                                    ckertype = 'gaussian',
                                    bws = 0.6)
)
exer_newdata <- data.frame(Alcohol.consumption = with(datajitt, 
                                                    seq(range(Alcohol.consumption)[1], range(Alcohol.consumption)[2], by = 0.01))
)
summary(m_loc_alc_sleep_gaus)
preds_alc_gaus <- predict(m_loc_alc_sleep_gaus, newdata = exer_newdata, se = TRUE)
se_bands_alc_gaus <- cbind(preds_alc_gaus$fit + 2 * preds_alc_gaus$se.fit, preds_alc_gaus$fit - 2 * preds_alc_gaus$se.fit)

m_loess_alc <- loess(Sleep.efficiency ~ Alcohol.consumption,
                      span = 0.35, 
                      degree = 0,  # Estimatore di Nadaraya-Watson
                      data = datajitt)

summary(m_loc_deep_sleep_gaus)
summary(m_loess_alc)


preds_loess_alc <- predict(m_loess_alc, newdata = exer_newdata, se = TRUE)
se_bands_loess_alc <- cbind(preds_loess_alc$fit + 2 * preds_loess_alc$se.fit, preds_loess_alc$fit - 2 * preds_loess_alc$se.fit)


with(datajitt,
     plot(Alcohol.consumption, Sleep.efficiency, 
          cex = 1,
          col = "gray35",
          main = "Local Regression with Gaussian Kernel",
          xlab = "Alcohol Consumption",
          ylab = "Sleep Efficiency"
     )
)

lines(exer_newdata$Alcohol.consumption, preds_alc_gaus$fit, lwd = 2, col = "orangered", lty = 1)
matlines(exer_newdata$Alcohol.consumption, se_bands_alc_gaus, lwd = 1, col = "orangered", lty = 3)
# lines(exer_newdata$Alcohol.consumption, preds_loess_alc$fit, lwd = 2, col = "orangered", lty = 1)
# matlines(exer_newdata$Alcohol.consumption, se_bands_loess_alc, lwd = 1, col = "orangered", lty = 3)
# legend("topright", 
#        legend = c("Gaussian Kernel", "LOESS"), 
#        col = c("red4", "orangered"), 
#        lty = c(1, 1), 
#        lwd = 2, 
#        title = "Methods",
#        bty = "n"
# )

```

```{r}

# plot(datajitt$Sleep.efficiency, datajitt$Deep.sleep.percentage)
m_poly_eff_deep <- m_list_eff_deep[[8]]
m_kern_smooth_eff_deep <- with(datajitt, 
                               KernSmooth::locpoly(x = Deep.sleep.percentage, y = Sleep.efficiency, bandwidth = 5.8,
                                                   degree = 2, range.x = c(range(Deep.sleep.percentage)[1],
                                                                           range(Deep.sleep.percentage)[2]),
                                                   gridsize = nrow(deep_newdata))
)

m_loess_eff_deep <- loess(Sleep.efficiency ~ Deep.sleep.percentage, span = 0.35, degree = 2, data = datajitt)
preds_loess_eff_deep <- predict(m_loess_eff_deep, newdata = deep_sleep_grid, se = TRUE)
se_bands_loess_eff_deep <- cbind(preds_loess_eff_deep$fit + 2 * preds_loess_eff_deep$se.fit, preds_loess_eff_deep$fit - 2 * preds_loess_eff_deep$se.fit)
preds_eff_deep <- predict(m_poly_eff_deep, list(Deep.sleep.percentage = deep_sleep_grid), se.fit = TRUE)
se_bands_eff_deep <- cbind(preds_eff_deep$fit + 2 * preds_eff_deep$se.fit, preds_eff_deep$fit - 2 * preds_eff_deep$se.fit)
plot(datajitt$Deep.sleep.percentage, datajitt$Sleep.efficiency, 
     xlim = range(deep_sleep_grid), cex = 0.5, col = "darkgrey", 
     main = "Deep Sleep Percentage vs Sleep Efficiency", 
     xlab = "Deep Sleep Percentage", ylab = "Sleep Efficiency"
)
lines(deep_sleep_grid, preds_eff_deep$fit, lwd = 2, col = "blue", lty = 1) 
matlines(deep_sleep_grid, se_bands_eff_deep, lwd = 1, col = "blue", lty = 3)
lines(m_kern_smooth_eff_deep$x, m_kern_smooth_eff_deep$y, lwd = 2, col = "green4", lty = 1) 

lines(deep_sleep_grid, preds_loess_eff_deep$fit, lwd = 2, col = "orangered", lty = 1) 
matlines(deep_sleep_grid, se_bands_loess_eff_deep, lwd = 1, col = "orangered", lty = 3)
legend("topleft", 
       legend = c("Polynomial (Degree 8)", "Kernel Smooth (Degree 8)", "LOESS (Degree 2)"), 
       col = c("blue", "green4", "orangered"), 
       lty = c(1, 1, 1), 
       lwd = 2, 
       title = "Methods",
       bty = "n"
)

# plot(datajitt$Age, datajitt$Exercise.frequency)
m_poly_age_exer <- m_list_age_exer[[2]]
m_kern_smooth_age_exer <- with(datajitt, 
                               KernSmooth::locpoly(x = Age, y = Exercise.frequency, bandwidth = 6, degree = 1,
                                                   range.x = c(range(Age)[1], range(Age)[2]), gridsize = nrow(age_newdata))
)
se_bands_kern_age_exer <- cbind(m_kern_smooth_age_exer$y + 2 * m_kern_smooth_age_exer$se, m_kern_smooth_age_exer$y - 2 * m_kern_smooth_age_exer$se)

m_loess_age_exer <- loess(Exercise.frequency ~ Age, span = 0.35, degree = 1, data = datajitt)
preds_loess_age_exer <- predict(m_loess_age_exer, newdata = age_grid, se = TRUE)
se_bands_loess_age_exer <- cbind(preds_loess_age_exer$fit + 2 * preds_loess_age_exer$se.fit, preds_loess_age_exer$fit - 2 * preds_loess_age_exer$se.fit)
preds_age_exer <- predict(m_poly_age_exer, list(Age = age_grid), se.fit = TRUE)
se_bands_age_exer <- cbind(preds_age_exer$fit + 2 * preds_age_exer$se.fit, preds_age_exer$fit - 2 * preds_age_exer$se.fit)
plot(datajitt$Age, datajitt$Exercise.frequency, 
     xlim = range(age_grid), cex = 0.5, col = "darkgrey", 
     main = "Age vs Exercise Frequency", 
     xlab = "Age", ylab = "Exercise Frequency"
)
lines(age_grid, preds_age_exer$fit, lwd = 2, col = "blue", lty = 1) 
matlines(age_grid, se_bands_age_exer, lwd = 1, col = "blue", lty = 3)
lines(m_kern_smooth_age_exer$x, m_kern_smooth_age_exer$y, lwd = 2, col = "green4", lty = 1) 

lines(age_grid, preds_loess_age_exer$fit, lwd = 2, col = "orangered", lty = 1) 
matlines(age_grid, se_bands_loess_age_exer, lwd = 1, col = "orangered", lty = 3)
legend("topleft", 
       legend = c("Polynomial (Degree 2)", "Kernel Smooth (Degree 1)", "LOESS (Degree 1)"), 
       col = c("blue", "green4", "orangered"), 
       lty = c(1, 1, 1), 
       lwd = 2, 
       title = "Methods",
       bty = "n"
)

# plot(datajitt$Awakenings, datajitt$Exercise.frequency)
m_poly_awak_exer <- m_list_exer_awak[[1]]

m_kern_smooth_awak_exer <- with(datajitt, 
                                KernSmooth::locpoly(x = Exercise.frequency, y = Awakenings, bandwidth = 0.6, degree = 1,
                                                    range.x = c(range(Exercise.frequency)[1], range(Exercise.frequency)[2]), gridsize = nrow(awak_newdata))
)

m_loess_awak_exer <- loess(Awakenings ~ Exercise.frequency, span = 0.35, degree = 1, data = datajitt)
preds_loess_awak_exer <- predict(m_loess_awak_exer, newdata = exer_grid, se = TRUE)
se_bands_loess_awak_exer <- cbind(preds_loess_awak_exer$fit + 2 * preds_loess_awak_exer$se.fit, preds_loess_awak_exer$fit - 2 * preds_loess_awak_exer$se.fit)

preds_awak_exer <- predict(m_poly_awak_exer, list(Exercise.frequency = exer_grid), se.fit = TRUE)
se_bands_awak_exer <- cbind(preds_awak_exer$fit + 2 * preds_awak_exer$se.fit, preds_awak_exer$fit - 2 * preds_awak_exer$se.fit)

plot(datajitt$Exercise.frequency, datajitt$Awakenings, 
     xlim = range(exer_grid), cex = 0.5, col = "darkgrey", 
     main = "Awakenings vs Exercise Frequency", 
     xlab = "Exercise Frequency", ylab = "Awakenings"
)
lines(exer_grid, preds_awak_exer$fit, lwd = 2, col = "blue", lty = 1) # Polinomiale
matlines(exer_grid, se_bands_awak_exer, lwd = 1, col = "blue", lty = 3)

lines(m_kern_smooth_awak_exer$x, m_kern_smooth_awak_exer$y, lwd = 2, col = "green4", lty = 1) # Kernel Smooth

lines(exer_grid, preds_loess_awak_exer$fit, lwd = 2, col = "orangered", lty = 1) # LOESS
matlines(exer_grid, se_bands_loess_awak_exer, lwd = 1, col = "orangered", lty = 3)

legend("topleft", 
       legend = c("Polynomial (Degree 1)", "Kernel Smooth (Degree 1)", "LOESS (Degree 1)"), 
       col = c("blue", "green4", "orangered"), 
       lty = c(1, 1, 1), 
       lwd = 2, 
       title = "Methods",
       bty = "n"
)


# plot(datajitt$Exercise.frequency, datajitt$Sleep.efficiency)

m_poly_exer_sleep <- m_list_exer_sleep[[1]] # Modello polinomiale di grado 1

m_kern_smooth_exer_sleep <- with(datajitt, 
                                 KernSmooth::locpoly(x = Exercise.frequency, y = Sleep.efficiency, bandwidth = 0.6, degree = 1,
                                                     range.x = c(range(Exercise.frequency)[1], range(Exercise.frequency)[2]), gridsize = nrow(exer_newdata))
)

m_loess_exer_sleep <- loess(Sleep.efficiency ~ Exercise.frequency, span = 0.35, degree = 1, data = datajitt)
preds_loess_exer_sleep <- predict(m_loess_exer_sleep, newdata = exer_grid, se = TRUE)
se_bands_loess_exer_sleep <- cbind(preds_loess_exer_sleep$fit + 2 * preds_loess_exer_sleep$se.fit, preds_loess_exer_sleep$fit - 2 * preds_loess_exer_sleep$se.fit)

preds_exer_sleep <- predict(m_poly_exer_sleep, list(Exercise.frequency = exer_grid), se.fit = TRUE)
se_bands_exer_sleep <- cbind(preds_exer_sleep$fit + 2 * preds_exer_sleep$se.fit, preds_exer_sleep$fit - 2 * preds_exer_sleep$se.fit)

plot(datajitt$Exercise.frequency, datajitt$Sleep.efficiency, 
     xlim = range(exer_grid), cex = 0.5, col = "darkgrey", 
     main = "Exercise Frequency vs Sleep Efficiency", 
     xlab = "Exercise Frequency", ylab = "Sleep Efficiency"
)
lines(exer_grid, preds_exer_sleep$fit, lwd = 2, col = "blue", lty = 1) # Polinomiale
matlines(exer_grid, se_bands_exer_sleep, lwd = 1, col = "blue", lty = 3)

lines(m_kern_smooth_exer_sleep$x, m_kern_smooth_exer_sleep$y, lwd = 2, col = "green4", lty = 1) # Kernel Smooth

lines(exer_grid, preds_loess_exer_sleep$fit, lwd = 2, col = "orangered", lty = 1) # LOESS
matlines(exer_grid, se_bands_loess_exer_sleep, lwd = 1, col = "orangered", lty = 3)

legend("topleft", 
       legend = c("Polynomial (Degree 1)", "Kernel Smooth (Degree 1)", "LOESS (Degree 1)"), 
       col = c("blue", "green4", "orangered"), 
       lty = c(1, 1, 1), 
       lwd = 2, 
       title = "Methods",
       bty = "n")

```

```{r}
with(datajitt, scatterplotMatrix(data.frame(Sleep.efficiency, Deep.sleep.percentage, Exercise.frequency, Age)))
```

##### bootstrap regression

```{r}

names(data)
# Y <- data$Sleep.efficiency
# x1 <- data$Sleep.duration
# x2 <- data$Bedtime
# x3 <- data$Age
# x4 <- data$Caffeine.consumption
# x5 <- data$Exercise.frequency
# x6 <- data$Alcohol.consumption
# x7 <- data$Deep.sleep.percentage
# x8 <- data$Awakenings
# x9 <- data$REM.sleep.percentage
# x10 <- data$Smoking.status

model_lmtest <- lm(Sleep.efficiency ~ .-AgeGroup, data = datajitt)
summary(model_lmtest)

# lm(Y ~ x1+x2+x3+x5+x7+x6+x10)
model_lm1 <- lm(Sleep.efficiency ~ Age+Exercise.frequency+ Alcohol.consumption+Deep.sleep.percentage+Smoking.status,
               data = datajitt)
summary(model_lm1)

model_lm2 <- lm(Sleep.efficiency ~ Age+Deep.sleep.percentage,
               data = datajitt) 
summary(model_lm2)

age.grid <- seq(min(datajitt$Age), max(datajitt$Age), length.out = 800)
deep.grid <- seq(min(datajitt$Deep.sleep.percentage), max(datajitt$Deep.sleep.percentage), length.out = 800)
grid <- expand.grid(Age = age.grid, `Deep.sleep.percentage` = deep.grid)

pred <- predict(model_lm2, newdata = grid)

library(rgl)
persp3d(age.grid, deep.grid, matrix(pred, nrow = length(age.grid)), 
        col = 'blue', border = "black", lwd = 0.3, xlab = "Age", ylab = "Deep Sleep (%)", zlab = "Sleep Efficiency")
with(datajitt, points3d(Age, Deep.sleep.percentage, Sleep.efficiency, col = 'black', size = 5))

model_lm_int <- lm(Sleep.efficiency ~ Exercise.frequency+Deep.sleep.percentage+Exercise.frequency:Deep.sleep.percentage,
                   data = datajitt) 
summary(model_lm_int)

pred_interaction <- predict(model_lm_int, newdata = grid)

exe.grid <- seq(min(datajitt$Exercise.frequency), max(datajitt$Exercise.frequency), length.out = 800)

{
   persp3d(exe.grid, deep.grid, matrix(pred_interaction, nrow = length(exe.grid)), 
        col = 'grey30', xlab = "Exercise.frequency", ylab = "Deep Sleep (%)", zlab = "Sleep Efficiency")
   with(datajitt, points3d(Exercise.frequency, Deep.sleep.percentage, Sleep.efficiency, col = 'black', size = 5))
}


griglia <- seq(0, 1)
griglia <- seq(0, 2500)
data$Sleep.efficiency
griglia <- seq(0, 1, 1000)
fitted.obs <- predict(model_lm1)
fitted.obs
res.obs <- Y - fitted.obs
plot(fitted.obs, Y)
hist(data$Sleep.efficiency)
hist(Y)
hist(datastd$Sleep.efficiency)
plot(fitted.obs, Y)
plot(model)
library(progress)
summary(model)$parameters[1, 1]
summary(model)$parameters[1, 1]
summary(model)$parameters[1, 1]
summary(model)$parameters
summary(model)
summary(model)$coefficients[1, 1]
L.obs <- summary(model)$coefficients[2, 1]
T.boot.L <- numeric(B)
B<-2000
seed<-156
L.obs <- summary(model)$coefficients[2, 1]
T.boot.L <- numeric(B)
formula.b <- response.b ~ logistic(age, L, k, midpoint)
formula.b <- responseage ~ x3+x5+x7+x6+x10
pb <- progress_bar$new(
format = "  processing [:bar] :percent eta: :eta",
total = B,
clear = FALSE)
set.seed(seed)
for (b in 1:B) {
response.b <- fitted.obs + sample(res.obs, replace = T)
fm.b <- lm(formula.b)
sm.b <- summary(fm.b)
T.boot.L[b] <- sm.b$coeff[2, 1]
pb$tick()
}
formula.b
for (b in 1:B) {
responseage <- fitted.obs + sample(res.obs, replace = T)
fm.b <- lm(formula.b)
sm.b <- summary(fm.b)
T.boot.L[b] <- sm.b$coeff[2, 1]
pb$tick()
}
hist(T.boot.L, main='Asymptote Bootstrap p.m.f.')
abline(v=L.obs, lty=2)
plot(ecdf(T.boot.L), main='Asymptote Bootstrap cdf')
abline(v=L.obs, lty=2)
sd(T.boot.L)
bias <- mean(T.boot.L) - L.obs
MSE <- var(T.boot.L) + (bias)^2
alpha <- 0.05
right.quantile.L <- quantile(T.boot.L, 1 - alpha/2)
left.quantile.L <- quantile(T.boot.L, alpha/2)
CI.RP.L <- c(L.obs - (right.quantile.L - L.obs),
L.obs,
L.obs - (left.quantile.L - L.obs))
names(CI.RP.L) <- c("lwr", "lvl", "upr")
CI.RP.L
attach(O_3)
plot(ecdf(T.boot.L), main = "L")
abline(v = L.obs, lty = 2)
abline(v = CI.RP.L)
plot(age, circumference, main = "Fitted curve vs data",
xlim = c(0, 2500), ylim = c(20, 200))
lines(age_fine, circumference_fine, col = "red")
abline(h = L.obs, lty = 2)
abline(h = CI.RP.L)
lines
## inter 0-3 fiorentina , saluti dai Gobbi
```
