
##### Initial cleaning
```{r}
rm(list=ls())
graphics.off()
remove(list = setdiff(ls(), lsf.str()))
cat('\14')
```

##### Libraries and packages
```{r}
library(stringi)
library(naniar)
library(ggplot2)
library(DepthProc)
library(RColorBrewer)
library(tidyverse)
library(dplyr)
library(corrplot)
```

##### Upload of dataset
```{r}
setwd(getwd())
data<-read.csv("Sleep_efficiency.csv")
n<-dim(data)[1]
p<-dim(data)[2]
names(data)
```

###### Convert variable "Smoking status" into a binary variable
```{r}
sum(is.na(data$Smoking.status)) # no NAs
data$Smoking.status <- as.factor(ifelse(data$Smoking.status == "Yes", 1, 0))
```

###### Convert Bedtime and Wakeup time
```{r}
data$Bedtime <- as.POSIXct(data$Bedtime, format = "%Y-%m-%d %H:%M:%S")
data$Wakeup.time <- as.POSIXct(data$Wakeup.time, format = "%Y-%m-%d %H:%M:%S")
# Estract the hour
data$Bedtime <- as.numeric(format(data$Bedtime, "%H"))
data$Bedtime <- ifelse(data$Bedtime < 4, data$Bedtime + 24, data$Bedtime)
data$Wakeup.time <- as.numeric(format(data$Wakeup.time, "%H"))
```

###### Grouping age by decades
```{r}
# sum(is.na(data$Age)) # no NAs
data$AgeGroup <- cut(data$Age, breaks = seq(min(data$Age), max(data$Age), 
                length.out = nclass.FD(data$Age)), include.lowest = T, right = T) 

table(data$AgeGroup, data$Gender)
```

##### NAs 
```{r}
summary(data) # 65 NAs 
sum(complete.cases(data)) # 388 complete observations, 64 incomplete
sort(colSums(is.na(data))[!(colSums(is.na(data))==0)],decreasing = T) 
```

###### NA imputation with Age_group median
```{r}
data <-data %>%
  group_by(AgeGroup, Gender) %>%        
  mutate(across(where(is.numeric),        
          ~ ifelse(is.na(.), median(., na.rm = TRUE), .))) %>%
  ungroup()     
summary(data)
```

###### Exploratory analysis: errors in Awakenings imputation
```{r}
levels(as.factor(data$Awakenings))
data$Awakenings<-round(data$Awakenings)
data$Exercise.frequency<-round(data$Exercise.frequency)
```

##### Correlazione
```{r}
numeric_cols <- which(sapply(data, is.numeric))[-1]
# pairs(data[,numeric_cols])
corr_matrix <- cor(data[,numeric_cols], method = "pearson")
# View(corr_matrix)
# heatmap(corr_matrix)
par(mfrow=c(1,1))
# corrplot(corr_matrix)
```

##### Standardize numeric variables 
```{r}
datastd<- scale(data[,numeric_cols])
datastd<-cbind(datastd,data[,-numeric_cols])
```

##### Split the dataset 
```{r}
datafemale <- filter(data, Gender == "Female")
datayoung <- filter(data, AgeGroup == "[16,23)")
datacoffee <- filter(data, Caffeine.consumption < 24)
rm(datafemale, datayoung, datacoffee)
```

##### Extract specific columns 
```{r}
sleepdata<-dplyr::select(data, Age, Gender)
rm(sleepdata)
```   

##### Data jittering 
```{r}
set.seed(1524)
datajitt <- data %>%
  mutate(across(c(6, 11, 12, 13, 14, 15), ~  pmax(0,jitter(as.numeric(.), length(levels(as.factor(.))),0)) ))
```

